Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-10-01 packages/spec-haskell/yellowpaper/Paper1.tex
refine payment primitives texts

One of the approaches is \textit{denotational semantics}, which is used in Chapter 3 of the paper to define the general
payment primitives. Along with the denotational semantics, a restatement\footnote{It is a borrowed term from common law:
"restatement of the law". In our case, the denotative mathematical "laws".} of it in \textit{Haskell programming
    language} (\cite{hudak1992report} \cite{jones2003haskell} \cite{marlow2010haskell}) is also included.

@@

\begin{itemize}
\item \textit{money distribution} models how monetary value is distributed amongst bearers\footnote{(Banking \& Finance)
a person who presents a note or bill for payment. - Collins English Dictionary},
\item \textit{payment primitives} updates money distribution,
\item \textit{payment execution environment} performs payment primitives,
\item and \textit{forms of money medium} are the "user interfaces" of money for the bearers.
\end{itemize}

@@

For the purpose of this paper, the model of context is: $\gamma = (t :: Timestamp) \oplus (ctx :: SharedContext)$.

@@

Payment primitives are functions that takes a money distribution, shared context, some payment instruction (args) and
timestamp as inputs and incremental updates\footnote{Specifically being monoidal, that is in short a set that has
associative binary operation and an identity element. See https://ncatlab.org/nlab/show/monoid. } of money
distribution and shared context as outputs:

\begin{minted}[escapeinside=||,mathescape=true]{haskell}
    somePaymentPrimitive :: ( Timestamp t
                            , MoneyDistribution md
                            , SharedContext ctx
                            , PaymentInstruction args
                            , Monoid md, Monoid ctx
                            )
                         => md |$\times$| ctx -> args -> t -> md |$\times$| ctx
\end{minted}

@@

@@

\begin{itemize}
\item The meaning of $[\![\mathcal{M}]\!]$ is money distribution.
\item Semantic function $[\![.]\!] :: \mathcal{M} \rightarrow [\![\mathcal{M}]\!] $ include payment primitives.
\end{itemize}

@@

A payment primitive \textit{prim} is a semantic function that produces a monoidal money distribution, semantically
representing an ``update'' to a money distribution:

\begin{equation}
    [\![prim\ args]\!] = [\![\Delta\ m]\!]
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\textit{law of conservation of value for payment primitives} \\
        &\forall t \in \mathbb{R} ({\displaystyle \sum_{u \in U} [\![prim\ args]\!]\ u\ t = 0}) \\
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\textit{money distribution consists of updates from payment primitives} \\
        &[\![m]\!] = [\![prim_1\ args_1]\!] \oplus [\![prim_2\ args_2]\!] \oplus [\![prim_3\ args_3]\!] \oplus \dotsb
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\textbf{In addition to the axioms, given:} \\
        \\
        &\textit{(base case with empty money distribution is trivial)} \\
        &\forall t \in \mathbb{R} {\displaystyle \sum_{u \in U} [\![\emptyset]\!]\ u\ t = 0} \\
        \\
        &\textit{(monoid homomorphism)} \\
        &{\displaystyle \sum_{u \in U} [\![prim_1\ args_1\oplus prim_2\ args_2]\!]\ u\ t} = \\
        &\qquad {\displaystyle \sum_{u \in U} [\![prim_2\ args_2]\!]\ u\ t} \oplus
          {\displaystyle \sum_{u \in U} [\![prim_2\ args_2]\!]\ u\ t} \\
        \\
        &\textbf{We have:} \\
        &\forall t \in \mathbb{R}. \\
        &{\displaystyle \sum_{u \in U} [\![m]\!]\ u\ t} \\
        &\textit{(applying Axiom B)} \\
        = &{\displaystyle \sum_{u \in U}
            ([\![prim_1\ args_1]\!] \oplus
            [\![prim_2\ args_2]\!] \oplus
            [\![prim_3\ args_3]\!]\oplus \dotsb)
        }\ u\ t\\
        &\textit{(applying monoid homomorphism)} \\
        = &{\displaystyle \sum_{u \in U} [\![prim_1\ args_1]\!]\ u\ t} \oplus
        {\displaystyle \sum_{u \in U} [\![prim_2\ args_2]\!]\ u\ t} \oplus
        {\displaystyle \sum_{u \in U} [\![prim_3\ args_3]\!]\ u\ t} \oplus \dotsb
        )\\
        &\textit{(applying Axiom A)} \\
        =&\ 0
        \\
        \blacksquare
    \end{split}
\end{equation}

@@

\section{Sub Systems}

\section{Agreement Sub System}

\section{Type of Agreements}

\subsection{Transferable Balance Agreement}

@@

\section{Solvency Sub Systems}

\subsection{Buffer Based Solvency Systems}

\section{Type Of Monetary Units}

>>>>

One of the approaches is \textit{denotational semantics}, which is used in Chapter 3 of the paper to define the general
payment primitives. Along with the denotational semantics, a restatement\footnote{It is a borrowed term from common law:
``restatement of the law''. In our case, the denotative mathematical laws.} of it in \textit{Haskell programming
    language} (\cite{hudak1992report} \cite{jones2003haskell} \cite{marlow2010haskell}) is also included.

@@

\begin{itemize}
\item \textit{money distribution} models how monetary value is distributed amongst bearers\footnote{(Banking \& Finance)
a person who presents a note or bill for payment. - Collins English Dictionary},
\item \textit{payment primitives} updates money distribution,
\item \textit{payment execution environment} performs payment primitives,
\item and \textit{forms of money medium} are the ``user interfaces'' of money for the bearers.
\end{itemize}

@@

For the purpose of this paper, the model of context is: $\gamma = t \times ctx$.

@@

A payment primitive is a data type with a generator that creates payment primitive data from the shared context $ctx$,
primitive specific argument $args$ and timestamp, along with an update of the shared context. Here is an example of
generator for payment primitive \textit{a}:

\begin{equation}
    genPrim_a :: ctx \rightarrow args_a \rightarrow t \rightarrow prim \times ctx
\end{equation}

Payment primitive data then can be used to create a delta update of money distribution\footnote{Specifically being
monoidal, that is in short a set that has associative binary operation and an identity element. See
https://ncatlab.org/nlab/show/monoid.}

\begin{equation}
    runPrim :: prim \rightarrow md
\end{equation}

@@

\subsection{$\mathcal{M}$ - Category of Money Distribution}

\begin{itemize}
\item The meaning of the mathematical \textit{model} $[\![\mathcal{M}]\!]$ is money distribution.
\item Semantic function $[\![.]\!] :: \mathcal{M} \rightarrow [\![\mathcal{M}]\!] $ evaluates the syntax of money
      distribution, payment primitives, etc.
\end{itemize}

@@

@@

A payment primitive \textit{prim} is a model that produces a monoidal money distribution, semantically
representing an ``update'' to a money distribution:

\begin{equation}
    [\![prim]\!] = [\![\Delta\ m]\!]
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\textit{law of conservation of value for payment primitives} \\
        &\forall t \in \mathbb{R} ({\displaystyle \sum_{u \in U} [\![prim]\!]\ u\ t = 0}) \\
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\textit{money distribution consists of updates from payment primitives} \\
        &[\![m]\!] = [\![prim_1]\!] \oplus [\![prim_2]\!] \oplus [\![prim_3]\!] \oplus \dotsb
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\textbf{In addition to the axioms, given:} \\
        \\
        &\textit{(base case with empty money distribution is trivial)} \\
        &\forall t \in \mathbb{R} {\displaystyle \sum_{u \in U} [\![\emptyset]\!]\ u\ t = 0} \\
        \\
        &\textit{(monoid homomorphism)} \\
        &{\displaystyle \sum_{u \in U} [\![prim_1 \oplus prim_2]\!]\ u\ t} =
          {\displaystyle \sum_{u \in U} [\![prim_2]\!]\ u\ t} \oplus
          {\displaystyle \sum_{u \in U} [\![prim_2]\!]\ u\ t} \\
        \\
        &\textbf{We have:} \\
        &\forall t \in \mathbb{R}. \\
        &{\displaystyle \sum_{u \in U} [\![m]\!]\ u\ t} \\
        &\textit{(applying Axiom B)} \\
        = &{\displaystyle \sum_{u \in U}
            ([\![prim_1]\!] \oplus
            [\![prim_2]\!] \oplus
            [\![prim_3]\!]\oplus \dotsb)
        }\ u\ t
        \\
        &\textit{(applying monoid homomorphism)} \\
        = &{\displaystyle \sum_{u \in U} [\![prim_1]\!]\ u\ t} \oplus
        {\displaystyle \sum_{u \in U} [\![prim_2]\!]\ u\ t} \oplus
        {\displaystyle \sum_{u \in U} [\![prim_3]\!]\ u\ t} \oplus \dotsb
        \\
        &\textit{(applying Axiom A)} \\
        =&\ 0
        \\
        \blacksquare
    \end{split}
\end{equation}

@@

\section{Agreement Framework}

\section{Hierarchy of Agreements}

\subsection{Instant Transferable Agreement}

@@
